---
title: "Data I/O Lab, Part 1"
output: html_document
---

```{r, include = TRUE}
library(tidyverse)
library(readxl)
```

This lab uses different versions of the `iris` dataset. You can use links or download them to your machine. Copy the code produced by the dropdown menu into the code chunk. As a reminder the steps are:

-  `>` File
-  `>` Import Dataset
-  `>` From Text (`readr`)
-  `>` paste the url
-  `>` click "Update" and "Import"

You can see more about this dataset by running `?iris` in the console.

If this is review to you, explore the "Import Options" on the bottom left of the import dataset menu.

1. Read in the `https://sisbid.github.io/Data-Wrangling/data/iris/iris_q1.csv` dataset into the `iris_q1` R object. What delimiter separates columns?

```{r}

```

2. Read in the `https://sisbid.github.io/Data-Wrangling/data/iris/iris_q2.tsv` dataset into the `iris_q2` R object. What delimiter separates columns?

```{r}

```

3. Read in the `https://sisbid.github.io/Data-Wrangling/data/iris/iris_q3.txt` dataset into the `iris_q3` R object. What delimiter separates columns?

```{r}

```

4. Read in the `https://sisbid.github.io/Data-Wrangling/data/iris/iris_q4.csv` dataset into the `iris_q4` R object. What delimiter separates columns? [hint: file extension is ambiguous]

```{r}

```

5. Read in the `https://sisbid.github.io/Data-Wrangling/data/iris/iris_xl.xlsx` dataset into the `iris_xl` R object. 

```{r}

```
